---
title: "ds2003_final_project_driscoll"
author: "fpd4fv"
date: '2022-12-01'
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(forcats)

library(ggplot2)
library(shiny)
library(plyr)

library(gridExtra)

```


Islamic education	9	Fiqh, Fiqh Fatwa, Usuluddin, Islamic education, Pendidikan Islam, irkhs, KIRKHS

Comp Science	17	BCS, IT, CTS, mathematics

Law	3	laws, law

Engineering		engin, engine, Engineering, ENM, KOE, KOP

Pyschology		psychology (3)

Business	2	Accounting, Banking Studies, BIT (10), Business Administration, Communication, econs, KENMS

Medicine		Biomedical science, Biotechnology, CTS, marine science, radiography, diploma nursing, nursing, MHSC

liberal arts		ALA, BENL (3), DIPLOMA TESL, Human Resources, human sciences, TAASL, malcom


```{r}
#reading in data, cleaning, and turning columns into factors

mhealth_data <- read.csv("Student_Mental_health.csv")

str(mhealth_data)
```


```{r}
### collapsing factors

#making everything in What.is.your.course lowercase to make it easier to collapse
mhealth_data <- mutate_all(mhealth_data, .funs=tolower)
View(mhealth_data)

#removing leading and trailing whitespace to prevent collapsing errors
mhealth_data$What.is.your.course. <- trimws(mhealth_data$What.is.your.course., which = c("both"))

#creating duplicate study column that will be collapsed into study categories
mhealth_data$study_category <- mhealth_data$What.is.your.course.

mhealth_data$study_category <- as.factor(mhealth_data$study_category)
summary(mhealth_data$study_category)


mhealth_data$study_category <- fct_collapse(mhealth_data$study_category,
  Islamic_education = c("fiqh", "fiqh fatwa", "usuluddin", "islamic education",
                        "pendidikan islam", "irkhs", "kirkhs"),
  comp_science = c("bcs", "it", "cts", "mathemathics"),
  Law	= c("laws", "law"),
  engineering = c("engin", "engine", "engineering", "enm", "koe", "kop"),
  psychology = c("psychology"),
  business = c("accounting", "banking studies", "bit", "business administration",
               "communication", "econs", "kenms"),
  medicine = c("biomedical science", "biotechnology", "cts", "marine science",
               "radiography", "diploma nursing", "nursing", "mhsc"),
  liberal_arts = c("ala", "benl","diploma tesl", "human resources", 
                   "human sciences", "taasl","malcom")
)


summary(mhealth_data$study_category)
#study categories have been collapsed into 8 factor levels


#making other columns factors as needed
to_factor_cols <- c("Choose.your.gender","Your.current.year.of.Study",
                    "What.is.your.CGPA.",
                    "Marital.status","Do.you.have.Depression.","Do.you.have.Anxiety.","Do.you.have.Panic.attack.","Did.you.seek.any.specialist.for.a.treatment.")
mhealth_data[to_factor_cols] <- lapply(mhealth_data[to_factor_cols], factor)

#making age an integer
mhealth_data$Age <- as.numeric(mhealth_data$Age)


str(mhealth_data)
```

```{r}

age_sub <- mhealth_data[mhealth_data$Age >= 18 & mhealth_data$Age <= 24, ]
final_sub <- age_sub

# cs <- colSums(final_sub$Marital.status == "yes")

# ggplot(final_sub, aes(x =`What.is.your.CGPA.`, y="", fill = `Marital.status`)) + geom_bar() + coord_polar("y", start=0) + theme_void() 


# marStat <- count(final_sub$Marital.status)


# ggplot(final_sub, aes(x =`What.is.your.CGPA.`, y="", fill = `Marital.status`)) + geom_col() + coord_polar(theta = "y")

ggplot(final_sub, aes(x="", y="", fill = `Marital.status`)) + geom_bar(stat="identity", width=1) + coord_polar("y", start=0) + theme_void() 
# What.is.your.CGPA 0 - 1.99

count(final_sub$`What.is.your.CGPA`)

```


```{r}
print(levels((mhealth_data$`What.is.your.CGPA`)))
```


```{r}
# Create Shiny Application
ui <- fluidPage(
#  checkboxGroupInput("gpaGroup", label = h3("GPA ranges"), 
#        choices = list("0 - 1.99" = "0 - 1.99", "2.00 - 2.49" = "2.00 - 2.49", "2.50 - 2.99" = "2.50 - 2.99", "3.00 - 3.49" = "3.00 - 3.49", "3.50 - 4.00" = "3.50 - 4.00"),
#        selected = "0 - 1.99"),
    checkboxGroupInput("gpaGroup", label = h3("GPA ranges"), 
        choices = c(levels(mhealth_data$`What.is.your.CGPA`)),
        selected = c(levels(mhealth_data$`What.is.your.CGPA`))),
  
  sliderInput("range", label = h3("Age Range"), min = 18, 
        max = 24, value = c(18, 24)),
  radioButtons("married", label = h3("Marital Status"),
        choices = list("Yes" = "yes", "No" = "no"), 
        selected = "yes"),
  
  plotOutput("GPA_Plot"))

  server <- function(input, output){
    
    output$GPA_Plot <- renderPlot({
      age_sub <- mhealth_data[mhealth_data$Age >= input$range[1] & mhealth_data$Age <= input$range[2], ] 
      married_sub <- age_sub[age_sub$`Marital.status` == input$married, ]
      gpa_sub <- married_sub[married_sub$`What.is.your.CGPA` %in% input$gpaGroup,]
      final_sub <- gpa_sub
      print(nrow(final_sub))
      
      validate(
        need(nrow(final_sub) > 0, "No Data with Provided Selections Exists")
      )
      
      DepPlot <- ggplot(final_sub, aes(x="", y="", fill = `Do.you.have.Depression.`)) + geom_bar(stat="identity", width=1) + coord_polar("y", start=0) + theme_void() 
      AnxPlot <- ggplot(final_sub, aes(x="", y="", fill = `Do.you.have.Anxiety.`)) + geom_bar(stat="identity", width=1) + coord_polar("y", start=0) + theme_void() 
      PanPlot <- ggplot(final_sub, aes(x="", y="", fill = `Do.you.have.Panic.attack.`)) + geom_bar(stat="identity", width=1) + coord_polar("y", start=0) + theme_void()
      
      grid.arrange(DepPlot, AnxPlot, PanPlot, ncol=3)
      
    })}
  
shinyApp(ui = ui, server = server, options = list(height = 600)) 
```

